CUBE: An Information-optimized parallel Cosmological $N$-body Algorithm

### 2017-11-06 ms_caf/ms_caf.tex
conclusion

\subsection{Code overview}\label{ss.overview}
{\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done. {\tt CUBE} is written in Coarray Fortran, where Coarray features \tcb{replace} MPI communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran. We use this terminology in this paper.}. The algorithm is described in this language.

@@

The upper parts of the four panels show the fractional power spectrum deviations. \tcb{For both simulations and redshifts, the fractional power spectrum deviation is less than 1\% on scales of interest, where PM error do not dominate. If we just simply observe the scales where PM already contributes significant errors, there are differences between simulations.} Because x2v2 gives exact results as {\tt CUBEP3M} on all scales \tcb{(even including the scale where PM contribute significant error)} $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines. Similar to the results form Figure \ref{fig.dsp}, x1v2 (red) curves give better results than x2v1 (blue) and x1v1 (green), and x2v1 and x1v1 give nearly same results. Another trend is that higher redshifts and/or larger box sizes gives much better results. In the case of S2048S $z=0$ it shows more deviations near Nyquist frequency.

@@

\section{Discussion and conclusion}\label{s.discussion}
We present an parallel integer-based $N$-body algorithm. This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g. \cite{2017PhRvD..95d3501Y,2017ApJ...841L..29W,2017MNRAS.469.1968P}. It needs very low memory usage, approaching 6 byte per particle (6 bpp). The accuracy of this 6pp configuration is acceptable on most scales of our interest, e.g. BAO. The accuracy is also adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms.

We analyze the deviation of power spectrum and de-correlation caused by integer-1 based algorithms in two simulations S2048S and S2048L. They have comparable number of particles per computing as the TianNu simulation. S2048S focuses on mass resolution -- comparable to TianNu and TianZero simulations, whereas S2048L focuses on simulation volume and have high
er particle mass. In the power spectra and cross correlations, S2048L performs better, and we conclude that in high-$m_p$ cases, we can safely use x1v1 in cosmological simulations.

@@

Traditional $N$-body codes consumes considerable memory while doing a relatively light computations. {\tt CUBE} prominently reduces the memory usage in $N$-body simulations and approaches toward a better balance between memory and computation. The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g. MIC and GPUs. Optimizing the velocity storage may further improve the accuracy of x1v1, and whose effects on neutrino-LSS simulation is going to be discovered.

>>>>

\subsection{Code overview}\label{ss.overview}
{\tt CUBE} uses a 2-level PM force calculation, same as {\tt CUBEP3M}, However, in order to apply the integer based format to the $N$-body simulation, substantial structural changes need to be done. {\tt CUBE} is written in Coarray Fortran, where Coarray features \tcb{replace} MPI (Message Passing Interface) communications between computation nodes/images\footnote{Images are the concept of computing nodes or MPI tasks in Coarray Fortran. We use this terminology in this paper.}. The algorithm is described in this language.

@@

The upper parts of the four panels show the fractional power spectrum deviations. \tcb{For both simulations and redshifts, the fractional power spectrum deviation is less than 1\% on scales of interest, where PM error do not dominate. If we just simply observe the scales where PM already contributes significant errors, there are differences between simulations.} For x2v2, it gives exact results as {\tt CUBEP3M} on all scales \tcb{(even including the scale where PM contribute significant error)} $\Delta^2_{\rm x2v2}(k)/\Delta^2_0(k)-1=0$ for both simulations and redshifts, shown by the four black straight lines. Similar to the results form Figure \ref{fig.dsp}, x1v2 (red) curves give better results than x2v1 (blue) and x1v1 (green), and x2v1 and x1v1 give nearly same results. Another trend is that higher redshifts and/or larger box sizes gives much better results. In the case of S2048S $z=0$ it shows more deviations near Nyquist frequency.

@@

\tcb{To summarize from Figure \ref{fig.ccc}, in terms of either power spectrum deviation or cross correlation, the error induced by information optimization (even for x1v1) is lower than the error from the PM algorithm, and we can safely use x1v1 for most of the LSS studies. If one still wants to compare the results when PM already contributes significant error ($k\gtrsim 0.2\ k_{\rm Nyquist}$), errors are from nonlinear (lower $m_p$ and/or lower redshift) and are primarily from compressing the velocity information (x2v1 and x1v1).}

\section{Discussion and conclusion}\label{s.discussion}
We present an parallel, \tcb{information-optimized} $N$-body algorithm. This open-source code {\tt CUBE}, has recently been used in many studies of LSS, e.g. \cite{2017PhRvD..95d3501Y,2017ApJ...841L..29W,2017MNRAS.469.1968P}. It needs very low memory usage, approaching 6 byte per particle (6 bpp).

The accuracy of this code is adjustable in that we can choose 1-byte/2-byte integers separately for positions and velocities of particles. In the case of using 2-byte integers for both positions and velocities (``x2v2'', and memory can be 12 bpp), the algorithm gives exact results given by traditional $N$-body algorithms. \tcb{Note that the results are exactly same in that they not only produce same physical statistics of LSS, but also the same error (not physical) from the PM algorithm near Nyquist frequencies. In another word, the positions and velocities of each particle are exact. In practice, we only require that the errors from information optimization is much lower than the errors from PM algorithm. In Figure \ref{fig.ccc} we see that it is the case even for the most memory efficient configuration x1v1. This shows that in most LSS studies, when our scales of interest are smaller than $k\simeq 0.2\ k_{\rm Nyquist}$, six 1-byte fixed point numbers contains sufficient information of every $N$-body particle.}

@@

Traditional $N$-body codes consumes considerable memory while doing a relatively light computations. \tcb{{\tt CUBE} is designed to optimize the efficiency of information and the memory usage in $N$-body simulations. {\tt CUBE} is written in Coarray Fortran -- the concise Coarray features are used instead of complicated MPI, and the code itself is much more concise than {\tt CUBEP3M} for future maintenance and development.} The next steps are optimization of the code and adapting it for various kinds of heterogeneous computing systems, e.g. MIC and GPUs. Optimizing the velocity storage may further improve the accuracy of x1v1, and whose effects on neutrino-LSS simulation is going to be discovered.
